---
title: "Assignment #2"
author: "Moise Dzogolo"
format: html
editor: visual
---

In this project, I use the `gtrendsR` package to request Google Trends data.The script I have used is below. The same data can also be downloaded from the Google Trends website, but gtrends gives more details.

Indeed, through the website, each dataset of interest must be downloaded individually. For example, one must download a trend over time, then trends by country, etc. With `gtrendsR`, on the other hand, a single line of code returns different dataset, and the user may select the one they wish to use as the request returns a list of datasets.

Another difference between downloaded data and the one from `gtrendsR` is that the downloaded data shows the date and time for each observation. The date seems to be missing from the `gtrendsR` returned data.

```{r}
## Load library and run gtrends
pacman::p_load(gtrendsR)
HarrisTrumpElection = gtrends(c("Trump","Harris","election"), time = "all")
```

```{r}
HarrisTrumpElection_interest <- HarrisTrumpElection$interest_over_time
## Plot data
par(family="Georgia")
plot(HarrisTrumpElection_interest$hits, type="l")
```

```{r}
## Collect data by timeframe

gtrends(c("Harris", "Trump"), time = "now 1-H") # last hour
# gtrends(c("Harris", "Trump"), time = "today 1-m") # last 30 days
```

```{r}
## Collect data by country

tg_gb <- gtrends(c("immigrants"), geo = c("GB", "US"), time = "now 1-H") 

## Check country codes
 
data("countries")
```
